---
title: "Bee analysis test"
author: "Eddy Dowle"
date: "2025-05-15"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(geomorph)
#setwd("C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Sam_april/Forewing")
```

## Geomorph
Using Geomorph package to do a generalized procrustes analysis on the TPS file and run basic statistics/plots. Not very familiar with the package so Im only using a very small subset of it and Im not convinced that I have everything correct/done in the best way (its been 15 years since I did this). 

```{r gpaallpoints}
#bring in TPS file
bees<-readland.tps('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Forewing_file_tps_dig_15_05_2025_1_renamed.TPS',specID = 'imageID')
#col_bees<-read.csv('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/bee_colours.csv')
#generate a file of species names for each sample etc for plotting
Sample_names<-labels(bees)[[3]]
#Sample_names<-gsub('.*\\\\','',Sample_names)
species_list<-gsub('_.*','',Sample_names)
#species_list<-gsub('L.Kanap','L.kanap',species_list)
unique(species_list)
#do a generalixed procrustes analysis on TPS
beesgpa<-gpagen(bees)
#checking procrustes fit
#plot before procrustes
plot(bees)
#plot after procrustes
plot(beesgpa)
#the first point looks very dodgy
PCA<-gm.prcomp(beesgpa$coords)
plot(PCA)
#thats all just those varable ones on the first point
plotOutliers(beesgpa$coords)
```

## Dropping first point
That first point just isnt stable so think we can drop it form now on. 

```{r gpadropfirst}
bees_removefirst <- bees[-1,,] 
beesgpa<-gpagen(bees_removefirst)
#looks more stable
plot(beesgpa)
#plot outliers
plotOutliers(beesgpa$coords)
#run PCA analysis
PCA<-gm.prcomp(beesgpa$coords)
#plot(PCA)
#plot with legend
PCaxis<-as.data.frame(PCA$x[,1:2])
PCaxis<-rownames_to_column(PCaxis,var='ID')
PCaxis$bee_species<-gsub('_.*','',PCaxis$ID)

#set colours
col<-as.character(c("dodgerblue2","tan1","gold1","springgreen4","orchid4","palegreen","firebrick1","cyan1","hotpink"))
names(col)<-as.character(unique(PCaxis$bee_species))
col
#extract PC % from sdev
PC1_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[1,1]*100,digits=2)
PC2_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[2,1]*100,digits=2)
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=2)+
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#
```

## Within Leioproctus Leioproctus
Just wanted to have a quick look just within leioproctus leiopructus as I notice on the PCA most of the variaiton is just between Neso species and Leio species


```{r gpaleio}
#just looking with in leioproctus (plus monticola)
#fulvesencs and paahua are quite different
#redo procrustes just with Leioproctus leioproctus species (plus monitola)
bees_leioproctus <- bees[,,-c(11:15,31:35)] 
bees_leioproctus <- bees_leioproctus[-1,,] 
beesgpa_leio<-gpagen(bees_leioproctus)
Sample_names_leio<-labels(bees_leioproctus)[[3]]
Sample_names_leio<-gsub('.*\\\\','',Sample_names_leio)
species_list_leio<-gsub('_.*','',Sample_names_leio)
species_list_leio<-gsub('L.Kanap','L.kanap',species_list_leio)

unique(species_list_leio)

plot(beesgpa_leio)
#plot outliers
plotOutliers(beesgpa_leio$coords)
#run PCA analysis
PCA<-gm.prcomp(beesgpa_leio$coords)
#summary(PCA)
#plot(PCA)
#plot with legend
PCaxis<-as.data.frame(PCA$x[,1:2])
PCaxis<-rownames_to_column(PCaxis,var='ID')
PCaxis$bee_species<-gsub('_.*','',PCaxis$ID)

PC1_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[1,1]*100,digits=2)
PC2_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[2,1]*100,digits=2)

ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=2)+
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#
```

## Shape variation
Just plotting out the shape shift of each species

```{r shapevar}
#shape change species compared to average
ref<-mshape(beesgpa$coords)
species_list

#bring in links file for plotting
bee_links<-read.table('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Sam_april/Forewing/links_all_landmarks_remove1_sam.txt',sep=',',header=F)

#nesocoletes (note average is average across all Leioproctus sp. just grouped for veiwing)
#paahua
gpl.mn<-mshape(beesgpa$coords[,,31:35])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#fulv
gpl.mn<-mshape(beesgpa$coords[,,c(11:15)])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#leioproctus leioproctus
#mont
gpl.mn<-mshape(beesgpa$coords[,,2:6])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#bolt
gpl.mn<-mshape(beesgpa$coords[,,c(1,7:10)])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)


#huakiwi
gpl.mn<-mshape(beesgpa$coords[,,16:20])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#imat
gpl.mn<-mshape(beesgpa$coords[,,21:25])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#kanap
gpl.mn<-mshape(beesgpa$coords[,,26:30])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#pango
gpl.mn<-mshape(beesgpa$coords[,,36:40])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#vest
gpl.mn<-mshape(beesgpa$coords[,,41:45])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)
#
```

## Running some tests
Im not familar with geomorph so this is largely taken from the tutorial online

## Relationship between size and shape

```{R geomorphstats}
#check if there is a relationship between size and shape
#just having a look at raw csize
#kanap seem largest in general for forewing
size<-as.data.frame(beesgpa$Csize)
arrange(size,desc(size$`beesgpa$Csize`)) #%>% print(n=nrow(.))
gdf<-geomorph.data.frame(beesgpa,species=species_list)
fit.size<-procD.lm(coords~log(Csize), data = gdf, print.progress=F)
summary(fit.size)
#is a significant relationship between size and shape
#visiualise
#plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='PredLine',col=as.factor(species_list),pch=19)
#plot with legend etc
p<-plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='PredLine')
p<-cbind(as.data.frame(p$plot_args$x),as.data.frame(p$plot_args$y))
names(p)[names(p) == 'p$plot_args$x'] <- 'PredLine'
names(p)[names(p) == 'p$plot_args$y'] <- 'PC1'
p<-rownames_to_column(p,var='ID')
p$bee_species<-gsub('_.*','',p$ID)

ggplot(p,aes(x=PredLine,y=PC1,colour = bee_species))+
  geom_point(size=2) +
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x='Predictor',y='PC1 for fitted values')

#same thing here:
#plot(fit.size,type='regression',reg.type="PredLine", predictor =log(gdf$Csize))
#first principal component scores fitted values - PredLine

#regression scores-standardised regression shapes 
#plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='RegScore',col=as.factor(species_list),pch=19)
#plot with legend etc
p<-plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='RegScore')
p<-cbind(as.data.frame(p$plot_args$x),as.data.frame(p$plot_args$y))
names(p)[names(p) == 'p$plot_args$x'] <- 'PredLine'
names(p)[names(p) == 'V1'] <- 'RegScore'
p<-rownames_to_column(p,var='ID')
p$bee_species<-gsub('_.*','',p$ID)

ggplot(p,aes(x=PredLine,y=RegScore,colour = bee_species))+
  geom_point(size=2)+
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x='Predictor',y='Regression Score')

#partial least squares for relationship between shape and size
PLS<-two.b.pls(log(gdf$Csize),gdf$coords,print.progress=F)
PLS
#
```

## Relationship between size, shape and species
using Procrustes ANOVA/regression. Not convienced that I have this right and that this is the best way in regards to the posthoc pairwise tests. There is another function which I havent really explored that might be better for this. 

```{R geomorphstats2}
#look at the influence of shape, size and species
fit.common<-procD.lm(coords~log(Csize)+species,data=gdf)
summary(fit.common)
PW<-pairwise(fit.common,groups=species_list)
summary(PW)
#relationship between size and species
fit1<-procD.lm(log(gdf$Csize)~species_list)
summary(fit1)
PW<-pairwise(fit1,groups=species_list)
summary(PW)
#relationship between shape and species
fit2<-procD.lm(gdf$coords~species_list)
summary(fit2)
PW<-pairwise(fit2,groups=species_list)
summary(PW)
#interaction model - shape differ between species, while accounting for shape co varying with size
#not convinced I have this model right, I think its an over simplification
fit3<-procD.lm(gdf$coords ~log(gdf$Csize)*species_list)
summary(fit3)
#posthoc test between groups
PW<-pairwise(fit3,groups=species_list)
summary(PW)
#
```

## Hind wing analysis
Same but on hind wing

```{r gpaallpoints_hind}
#bring in TPS file
bees<-readland.tps('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Hindwings_merged_cleaned_19_05_2025.TPS',specID = 'imageID')

#generate a file of species names for each sample etc for plotting
Sample_names<-labels(bees)[[3]]
Sample_names
#Sample_names<-gsub('.*\\\\','',Sample_names)
#Sample_names<-sub("([A-Za-z].[A-Za-z]+_[0-9]+).*", "\\1", Sample_names)
species_list<-gsub('_.*','',Sample_names)
#species_list<-gsub('L.Kanap','L.kanap',species_list)
unique(species_list)
#do a generalixed procrustes analysis on TPS
beesgpa<-gpagen(bees)
#checking procrustes fit
#plot before procrustes
plot(bees)
#plot after procrustes
plot(beesgpa)
#plot outliers
plotOutliers(beesgpa$coords)
#the first point looks very dodgy
PCA<-gm.prcomp(beesgpa$coords)
plot(PCA)
#thats all just those varable ones on the first point again so doesnt work on either fore or hind wing for point 1
```

## Dropping first point
That first point just isnt stable so think we can drop it form now on. 

```{r gpadropfirst_hind}
bees_removefirst <- bees[-1,,] 
beesgpa<-gpagen(bees_removefirst)
#looks more stable
plot(beesgpa)
#plot outliers
plotOutliers(beesgpa$coords)
#run PCA analysis
PCA<-gm.prcomp(beesgpa$coords)
#plot(PCA)
#plot with legend
PCaxis<-as.data.frame(PCA$x[,1:2])
PCaxis<-rownames_to_column(PCaxis,var='ID')
PCaxis$bee_species<-gsub('_.*','',PCaxis$ID)

#set colours
col<-as.character(c("dodgerblue2","tan1","gold1","springgreen4","orchid4","palegreen","firebrick1","cyan1","hotpink"))
names(col)<-as.character(unique(PCaxis$bee_species))
col
#extract PC % from sdev
PC1_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[1,1]*100,digits=2)
PC2_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[2,1]*100,digits=2)
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=2)+
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#
```

## Within Leioproctus Leioproctus
Just wanted to have a quick look just within leioproctus leiopructus (plus monitcola) as I notice on the PCA most of the variaiton is just between two neso species and Leio species (plus monitcola)


```{r gpaleio_hind}
#just looking with in leioproctus
#I think fulvesencs and paahua are only neso but could be wrong
#redo procrustes just with Leioproctus leioproctus species
Sample_names
bees_leioproctus <- bees[,,-c(11:15,31:35)] 
bees_leioproctus <- bees_leioproctus[-1,,] 
beesgpa_leio<-gpagen(bees_leioproctus)
Sample_names_leio<-labels(bees_leioproctus)[[3]]
Sample_names_leio<-gsub('.*\\\\','',Sample_names_leio)
species_list_leio<-gsub('_.*','',Sample_names_leio)
species_list_leio<-gsub('L.Kanap','L.kanap',species_list_leio)

unique(species_list_leio)

plot(beesgpa_leio)
#plot outliers
plotOutliers(beesgpa_leio$coords)
#run PCA analysis
PCA<-gm.prcomp(beesgpa_leio$coords)
#summary(PCA)
#plot(PCA)
#plot with legend
PCaxis<-as.data.frame(PCA$x[,1:2])
PCaxis<-rownames_to_column(PCaxis,var='ID')
PCaxis$bee_species<-gsub('_.*','',PCaxis$ID)

PC1_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[1,1]*100,digits=2)
PC2_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[2,1]*100,digits=2)

ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=2)+
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#
```

## Shape variation
Just plotting out the shape shift of each species

```{r shapevar_hind}
#shape change species compared to average
ref<-mshape(beesgpa$coords)
species_list

#bring in links file for plotting
bee_links<-read.table('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Sam_april/Forewing/links_all_landmarks_remove1_sam_hind.txt',sep=',',header=F)

#nesocoletes (note average is average across all Leioproctus sp. just grouped for veiwing)
#paahua
gpl.mn<-mshape(beesgpa$coords[,,31:35])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#fulv
gpl.mn<-mshape(beesgpa$coords[,,c(11:15)])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#mont
gpl.mn<-mshape(beesgpa$coords[,,2:6])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#leioproctus leioproctus
#bolt
gpl.mn<-mshape(beesgpa$coords[,,c(1,7:10)])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#huakiwi
gpl.mn<-mshape(beesgpa$coords[,,16:20])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#imat
gpl.mn<-mshape(beesgpa$coords[,,21:25])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#kanap
gpl.mn<-mshape(beesgpa$coords[,,26:30])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#pango
gpl.mn<-mshape(beesgpa$coords[,,36:40])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)

#vest
gpl.mn<-mshape(beesgpa$coords[,,41:45])
#plotRefToTarget(M1=ref,M2=gpl.mn,mag=2)
plotRefToTarget(M1=ref,M2=gpl.mn,mag=2,links=bee_links)
#
```

## Running some tests
Im not familar with geomorph so this is largely taken from the tutorial online

## Relationship between size and shape

```{R geomorphstats_hind}
#check if there is a relationship between size and shape
#just having a look at raw csize
#size variation for hindwing
size<-as.data.frame(beesgpa$Csize)
arrange(size,desc(size$`beesgpa$Csize`)) #%>% print(n=nrow(.))
#turn that into a barchart when we have enough points per species
gdf<-geomorph.data.frame(beesgpa,species=species_list)
fit.size<-procD.lm(coords~log(Csize), data = gdf, print.progress=F)
summary(fit.size)
#is a significant relationship between size and shape
#visiualise
#plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='PredLine',col=as.factor(species_list),pch=19)
#plot with legend etc
p<-plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='PredLine')
p<-cbind(as.data.frame(p$plot_args$x),as.data.frame(p$plot_args$y))
names(p)[names(p) == 'p$plot_args$x'] <- 'PredLine'
names(p)[names(p) == 'p$plot_args$y'] <- 'PC1'
p<-rownames_to_column(p,var='ID')
p$bee_species<-gsub('_.*','',p$ID)

ggplot(p,aes(x=PredLine,y=PC1,colour = bee_species))+
  geom_point(size=2) +
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x='Predictor',y='PC1 for fitted values')

#same thing here:
#plot(fit.size,type='regression',reg.type="PredLine", predictor =log(gdf$Csize))
#first principal component scores fitted values - PredLine

#regression scores-standardised regression shapes 
#plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='RegScore',col=as.factor(species_list),pch=19)
#plot with legend etc
p<-plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='RegScore')
p<-cbind(as.data.frame(p$plot_args$x),as.data.frame(p$plot_args$y))
names(p)[names(p) == 'p$plot_args$x'] <- 'PredLine'
names(p)[names(p) == 'V1'] <- 'RegScore'
p<-rownames_to_column(p,var='ID')
p$bee_species<-gsub('_.*','',p$ID)

ggplot(p,aes(x=PredLine,y=RegScore,colour = bee_species))+
  geom_point(size=2)+
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x='Predictor',y='Regression Score')

#partial least squares for relationship between shape and size
PLS<-two.b.pls(log(gdf$Csize),gdf$coords,print.progress=F)
PLS
#
```

## Relationship between size, shape and species
using Procrustes ANOVA/regression. Not convienced that I have this right and that this is the best way in regards to the posthoc pairwise tests. There is another function which I havent really explored that might be better for this. 

```{R geomorphstats2_hind}
#look at the influence of shape, size and species
fit.common<-procD.lm(coords~log(Csize)+species,data=gdf)
summary(fit.common)
PW<-pairwise(fit.common,groups=species_list)
summary(PW)
#relationship between size and species
fit1<-procD.lm(log(gdf$Csize)~species_list)
summary(fit1)
PW<-pairwise(fit1,groups=species_list)
summary(PW)
#relationship between shape and species
fit2<-procD.lm(gdf$coords~species_list)
summary(fit2)
PW<-pairwise(fit2,groups=species_list)
summary(PW)
#interaction model - shape differ between species, while accounting for shape co varying with size
#not convinced I have this model right, I think its an over simplification
fit3<-procD.lm(gdf$coords ~log(gdf$Csize)*species_list)
summary(fit3)
#posthoc test between groups
PW<-pairwise(fit3,groups=species_list)
summary(PW)
#
```

There is bunch of other functions in geomorph, including morphol.dispartiy which I havent really figured out. I need to read the guide properly and relearn some geometrics. And there is another package morpho in R which I havent looked into. 

## Testing merging fore and hind wing

Just trying this out

```{R merge_foreandhind}
#forewing
bees_fore<-readland.tps('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Forewing_file_tps_dig_15_05_2025_1_renamed.TPS',specID = 'imageID')
Sample_names<-labels(bees_fore)[[3]]
species_list<-gsub('_.*','',Sample_names)
unique(species_list)
bees_fore_removefirst <- bees_fore[-1,,] 
beesgpa_fore<-gpagen(bees_fore_removefirst)
#hindwing
bees_hind<-readland.tps('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Hindwings_merged_cleaned_19_05_2025.TPS',specID = 'imageID')
Sample_names<-labels(bees_hind)[[3]]
species_list<-gsub('_.*','',Sample_names)
unique(species_list)
bees_hind_removefirst <- bees_hind[-1,,] 
beesgpa_hind<-gpagen(bees_hind_removefirst)
#combining
?combine.subsets
for_lab<-labels(beesgpa_fore$coords)[[3]]
hin_lab<-labels(beesgpa_hind$coords)[[3]]
match(for_lab,hin_lab)
#so ordered the same
#cant find a way to keep the names during the merge but I assume they add in by raw order? Might have to check this with package authors
#seems to be an active google groups page
#this combines with SCS or standard centroid size (but Im not sure if that is cs or relcs in the output)
comb.lm <- combine.subsets(forewing= beesgpa_fore, hindwing = beesgpa_hind, gpa = TRUE)
summary(comb.lm)
#just kinda plonks them on top of each other
plotAllSpecimens(comb.lm$coords)
#so it just plonks them on top of each other, visualised here with black and red
plot(comb.lm$coords[,,1], pch=21, bg = c(rep(1,18), rep(2,7)),asp =1,cex=1.5)
#
```

## Stats on combined data points

Im not familiar with this approach.
I found this which goes into it a bit https://github.com/DworkinLab/VirtualMorphoMeetup/blob/master/combine.subsets.talk/heads.or.tails.combined.landmarks.md
and suggests reading 
https://faculty.sites.iastate.edu/dcadams/files/inline-files/2020-evolbiol-collyer_et_al.pdf

"We recommend that empiricists neither perform GPA on combined confgurations, nor consider combined confgurations to be organismal “shape”. Rather, combined confgurations are shape composites which can be disjoined for the purpose of shape visualisation, but used in conjuction for statistical analyses'

So we cant draw out any changes but can analyse things statistically I think

https://onlinelibrary.wiley.com/doi/full/10.1002/jmor.21516
https://github.com/corinthiablack/Chew-on-this_Raw-data has some analyses with it

```{R merge_foreandhind_stats}
#run PCA analysis
PCA<-gm.prcomp(comb.lm$coords)
gm.prcomp(comb.lm$GPA)
#comb.lm$coords
#summary(PCA)
plot(PCA)
#plot with legend
PCaxis<-as.data.frame(PCA$x[,1:2])
#no names in data
PCaxis$ID<-for_lab
#PCaxis<-rownames_to_column(PCaxis,var='ID')
PCaxis$bee_species<-gsub('_.*','',PCaxis$ID)

PC1_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[1,1]*100,digits=2)
PC2_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[2,1]*100,digits=2)

ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=2)+
  theme_bw()+
  scale_color_manual(values=col)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#
```

## Stuck
Im not confident in the way to analyse the combined dataset

```{R geomorphstats_hindfor_merged}
#check if there is a relationship between size and shape
#just having a look at raw csize
#size variation for hindwing
size<-as.data.frame(comb.lm$CS)
#weird c-size came out with a name here, maybe its a bug that the coords are unnamed?
arrange(size,desc(size$forewing)) #%>% print(n=nrow(.))
#this doesnt work:
#gdf<-geomorph.data.frame(comb.lm,species=species_list)
#can bring it in seperatly like this
gdf<-geomorph.data.frame(Coords=comb.lm$coords,species=species_list)
#and even if it did I dont know how to deal with the allometry for the combined dataset as each sample has two csize values etc. So would it just be a shape ~species model  
fit<-procD.lm(Coords~species,data=gdf)
summary(fit)
PW<-pairwise(fit,groups=species_list)
summary(PW)

#can bring in cs and relcs seperatly too
gdf<-geomorph.data.frame(Coords=comb.lm$coords,relcs=comb.lm$rel.CS,cs=comb.lm$CS,species=species_list)

#But I dont know how to deal with the allometry for the combined dataset as each sample has two csize values etc. So would it just be a shape ~ species and size ~ species model?

#look at the influence of shape, size and species
#size and species using relative size
#I think this is SCS or standardised CS but Im not 100% sure its poorly described
fit1<-procD.lm(log(relcs)~species,data=gdf)
summary(fit1)
PW<-pairwise(fit1,groups=species_list)
summary(PW)
#size and species using raw cs
fit.common<-procD.lm(log(cs)~species,data=gdf)
summary(fit.common)
PW<-pairwise(fit.common,groups=species_list)
summary(PW)
#shape and species
fit<-procD.lm(Coords~species,data=gdf)
summary(fit)
PW<-pairwise(fit,groups=species_list)
summary(PW)
#cant do size and shape and species
#fit<-procD.lm(Coords~log(cs)+species,data=gdf)
#summary(fit)
#PW<-pairwise(fit,groups=species_list)
#summary(PW)

#I dont understand the difference between cs and relcs
#cant do size and shape and species
#fit<-procD.lm(Coords~log(relcs)+species,data=gdf)
#summary(fit)
#PW<-pairwise(fit,groups=species_list)
#summary(PW)
#
```
