---
title: "Bee analysis test hind"
author: "Eddy Dowle"
date: "2025-05-15"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(geomorph)
library(ggrepel)
#setwd("C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Sam_april/Forewing")
```

## Geomorph
Using Geomorph package to do a generalized procrustes analysis on the TPS file and run basic statistics/plots. Not very familiar with the package so Im only using a very small subset of it and Im not convinced that I have everything correct/done in the best way (its been 15 years since I did this). 

```{r gpaallpoints}
#bring in TPS file
bees<-readland.tps('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Hindwingd_from19_05_20_cleaned-27_05_25_ejd.TPS',specID = 'imageID',negNA = TRUE)
#col_bees<-read.csv('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/bee_colours.csv')
#generate a file of species names for each sample etc for plotting
Sample_names<-labels(bees)[[3]]
#Sample_names<-gsub('.*\\\\','',Sample_names)
species_list<-gsub('_.*','',Sample_names)
#species_list<-gsub('L.Kanap','L.kanap',species_list)
unique(species_list)
species_list_barry<-gsub('_[0-9].*','',Sample_names)
unique(species_list_barry)
#estimate missing
bees<-estimate.missing(bees, method = "TPS")
#do a generalixed procrustes analysis on TPS
beesgpa<-gpagen(bees)
#checking procrustes fit
#plot before procrustes
plot(bees)
#plot after procrustes
plot(beesgpa)
#first point looks lots better now was an issue with the missing data it turns out
PCA<-gm.prcomp(beesgpa$coords)
plot(PCA)
plotOutliers(beesgpa$coords)
#plot with legend
PCaxis<-as.data.frame(PCA$x[,1:2])
PCaxis<-rownames_to_column(PCaxis,var='ID')
PCaxis$bee_species<-gsub('_.*','',PCaxis$ID)

#set colours
as.character(unique(PCaxis$bee_species))
#18 species
#original colours
#col<-as.character(c("dodgerblue2","tan1","gold1","springgreen4","orchid4","palegreen","firebrick1","cyan1","hotpink"))
col<-as.character(c("dodgerblue2","tan1","gold1","springgreen4","orchid4","palegreen","firebrick1","cyan1","hotpink",'blue4',"lightpink","coral4","cadetblue","plum1","chartreuse1","lightgoldenrodyellow","grey","lightskyblue"))
names(col)<-as.character(c("L.bolt","L.mont","L.fulv","L.huak","L.imat","L.kanap","L.paahua","L.pango","Vest","L.hudsoni","L.hukare","L.keehua","L.marit","L.metal","L.pekanui","L.purp","L.unknown","L.waipou"))
#keeping same colour set as for forwing
#names(col)<-as.character(unique(PCaxis$bee_species))
col
#extract PC % from sdev
PC1_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[1,1]*100,digits=2)
PC2_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[2,1]*100,digits=2)
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species))+
  theme_bw()+
  scale_fill_manual(values=col)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#
#colouring barry's ones seperatly
#set colours
PCaxis$bee_species_baz<-gsub('_[0-9].*','',PCaxis$ID)
as.character(unique(PCaxis$bee_species))
as.character(unique(PCaxis$bee_species_baz))
#20 species
col_baz<-as.character(c("dodgerblue2","tan1","gold1","springgreen4","orchid4","palegreen","firebrick1","cyan1","hotpink",'blue4',"lightpink","coral4","tomato","cadetblue","plum1","burlywood","chartreuse1","lightgoldenrodyellow","grey","lightskyblue"))
names(col_baz)<-as.character(c("L.bolt","L.mont","L.fulv","L.huak","L.imat","L.kanap","L.paahua","L.pango","Vest","L.bolt_Barry","L.hudsoni_Barry","L.hukare_Barry", "L.keehua_Barry","L.marit_Barry","L.metal","L.pango_Barry","L.pekanui_Barry","L.purp_Barry","L.unknown","L.waipou_Barry"))
col_baz
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_baz)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')

#gods it confusing just comparing barrys to sams
#boltoni
col_one<-as.character(ifelse(unique(PCaxis$bee_species_baz) == 'L.bolt','red','grey'))
names(col_one)<-as.character(unique(PCaxis$bee_species_baz))
col_one<-replace(col_one, 'L.bolt_Barry', 'green')
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_one)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA boltoni")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')

#pango
col_one<-as.character(ifelse(unique(PCaxis$bee_species_baz) == 'L.pango','red','grey'))
names(col_one)<-as.character(unique(PCaxis$bee_species_baz))
col_one<-replace(col_one, 'L.pango_Barry', 'green')
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_one)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA pango")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#hukare
col_one<-as.character(ifelse(unique(PCaxis$bee_species_baz) == 'L.huak','red','grey'))
names(col_one)<-as.character(unique(PCaxis$bee_species_baz))
col_one<-replace(col_one, 'L.hukare_Barry', 'green')
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_one)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA huakare")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#Leioproctus leioproctus
col_leio<-col_baz
col_leio<-replace(col_leio, 'L.fulv', 'grey')
col_leio<-replace(col_leio, 'L.paahua', 'grey')
col_leio<-replace(col_leio, 'L.marit_Barry', 'grey')
col_leio<-replace(col_leio, 'L.pekanui_Barry', 'grey')
col_leio<-replace(col_leio, 'L.mont', 'grey')
col_leio<-replace(col_leio, 'L.hudsoni_Barry', 'grey')
#shifting some colours around to make it easier to see
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_leio)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA leio leio")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#leioproctus nesocolletes
col_neso<-col_baz
col_neso<-replace(col_neso, 'L.keehua_Barry', 'grey')
col_neso<-replace(col_neso, 'L.metal', 'grey')
col_neso<-replace(col_neso, 'L.imat', 'grey')
col_neso<-replace(col_neso, 'L.purp_Barry', 'grey')
col_neso<-replace(col_neso, 'Vest', 'grey')
col_neso<-replace(col_neso, 'L.kanap', 'grey')
col_neso<-replace(col_neso, 'L.waipou_Barry', 'grey')
col_neso<-replace(col_neso, 'L.bolt', 'grey')
col_neso<-replace(col_neso, 'L.bolt_Barry', 'grey')
col_neso<-replace(col_neso, 'L.pango', 'grey')
col_neso<-replace(col_neso, 'L.pango_Barry', 'grey')
col_neso<-replace(col_neso, 'L.huak', 'grey')
col_neso<-replace(col_neso, 'L.hukare_Barry', 'grey')
#shifting some colours around to make it easier to see
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_neso)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA leio nesi")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#having a look at sams unknowns
col_one<-as.character(ifelse(unique(PCaxis$bee_species_baz) == 'L.unknown','red','grey'))
#col_one
names(col_one)<-as.character(unique(PCaxis$bee_species_baz))
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_one)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA uknowns")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')+ 
  geom_text_repel(data=subset(PCaxis, bee_species_baz == 'L.unknown'),
            aes(x=Comp1,y=Comp2,label=ID))

#shape extremes in PCA
#bring in links file for plotting
bee_links<-read.table('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Sam_april/Forewing/links_all_landmarks_sam_hind.txt',sep=',',header=F)
#find mean shape
M<-mshape(beesgpa$coord)
#find extremes of PC1 (max/min)
PC<-PCA$x[,1]
preds<-shape.predictor(beesgpa$coords,x=PC,Intercept=F,pred1=min(PC),pred2=max(PC))
#PC1 min
plotRefToTarget(M,preds$pred1,links=bee_links)
#PC1 max
plotRefToTarget(M,preds$pred2,links=bee_links)
PC<-PCA$x[,2]
preds<-shape.predictor(beesgpa$coords,x=PC,Intercept=F,pred1=min(PC),pred2=max(PC))
#PC2 min
plotRefToTarget(M,preds$pred1,links=bee_links)
#PC2 max
plotRefToTarget(M,preds$pred2,links=bee_links)

```

## Dropping first point to match other bee analyses
It actually looks much better now with point 1 in so should do it with and without now

```{r gpadropfirst}
bees_removefirst <- bees[-1,,] 
beesgpa<-gpagen(bees_removefirst)
#looks more stable
plot(beesgpa)
#plot outliers
plotOutliers(beesgpa$coords)
#run PCA analysis
PCA<-gm.prcomp(beesgpa$coords)
#plot(PCA)
#plot with legend
PCaxis<-as.data.frame(PCA$x[,1:2])
PCaxis<-rownames_to_column(PCaxis,var='ID')
PCaxis$bee_species<-gsub('_.*','',PCaxis$ID)

#set colours
as.character(unique(PCaxis$bee_species))
#18 species
#original colours
#col<-as.character(c("dodgerblue2","tan1","gold1","springgreen4","orchid4","palegreen","firebrick1","cyan1","hotpink"))
col<-as.character(c("dodgerblue2","tan1","gold1","springgreen4","orchid4","palegreen","firebrick1","cyan1","hotpink",'blue4',"lightpink","coral4","cadetblue","plum1","chartreuse1","lightgoldenrodyellow","grey","lightskyblue"))
names(col)<-as.character(c("L.bolt","L.mont","L.fulv","L.huak","L.imat","L.kanap","L.paahua","L.pango","Vest","L.hudsoni","L.hukare","L.keehua","L.marit","L.metal","L.pekanui","L.purp","L.unknown","L.waipou"))
col
#extract PC % from sdev
PC1_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[1,1]*100,digits=2)
PC2_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[2,1]*100,digits=2)
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species))+
  theme_bw()+
  scale_fill_manual(values=col)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#
#colouring barry's ones seperatly
#set colours
PCaxis$bee_species_baz<-gsub('_[0-9].*','',PCaxis$ID)
as.character(unique(PCaxis$bee_species))
as.character(unique(PCaxis$bee_species_baz))
#20 species
col_baz<-as.character(c("dodgerblue2","tan1","gold1","springgreen4","orchid4","palegreen","firebrick1","cyan1","hotpink",'blue4',"lightpink","coral4","tomato","cadetblue","plum1","burlywood","chartreuse1","lightgoldenrodyellow","grey","lightskyblue"))
names(col_baz)<-as.character(c("L.bolt","L.mont","L.fulv","L.huak","L.imat","L.kanap","L.paahua","L.pango","Vest","L.bolt_Barry","L.hudsoni_Barry","L.hukare_Barry", "L.keehua_Barry","L.marit_Barry","L.metal","L.pango_Barry","L.pekanui_Barry","L.purp_Barry","L.unknown","L.waipou_Barry"))
col_baz
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_baz)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')

#gods it confusing just comparing barrys to sams
#boltoni
col_one<-as.character(ifelse(unique(PCaxis$bee_species_baz) == 'L.bolt','red','grey'))
names(col_one)<-as.character(unique(PCaxis$bee_species_baz))
col_one<-replace(col_one, 'L.bolt_Barry', 'green')
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_one)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA boltoni")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')

#pango
col_one<-as.character(ifelse(unique(PCaxis$bee_species_baz) == 'L.pango','red','grey'))
names(col_one)<-as.character(unique(PCaxis$bee_species_baz))
col_one<-replace(col_one, 'L.pango_Barry', 'green')
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_one)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA pango")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#hukare
col_one<-as.character(ifelse(unique(PCaxis$bee_species_baz) == 'L.huak','red','grey'))
names(col_one)<-as.character(unique(PCaxis$bee_species_baz))
col_one<-replace(col_one, 'L.hukare_Barry', 'green')
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_one)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA huakare")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#Leioproctus leioproctus
col_leio<-col_baz
col_leio<-replace(col_leio, 'L.fulv', 'grey')
col_leio<-replace(col_leio, 'L.paahua', 'grey')
col_leio<-replace(col_leio, 'L.marit_Barry', 'grey')
col_leio<-replace(col_leio, 'L.pekanui_Barry', 'grey')
col_leio<-replace(col_leio, 'L.mont', 'grey')
col_leio<-replace(col_leio, 'L.hudsoni_Barry', 'grey')
#shifting some colours around to make it easier to see
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_leio)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA leio leio")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#leioproctus nesocolletes
col_neso<-col_baz
col_neso<-replace(col_neso, 'L.keehua_Barry', 'grey')
col_neso<-replace(col_neso, 'L.metal', 'grey')
col_neso<-replace(col_neso, 'L.imat', 'grey')
col_neso<-replace(col_neso, 'L.purp_Barry', 'grey')
col_neso<-replace(col_neso, 'Vest', 'grey')
col_neso<-replace(col_neso, 'L.kanap', 'grey')
col_neso<-replace(col_neso, 'L.waipou_Barry', 'grey')
col_neso<-replace(col_neso, 'L.bolt', 'grey')
col_neso<-replace(col_neso, 'L.bolt_Barry', 'grey')
col_neso<-replace(col_neso, 'L.pango', 'grey')
col_neso<-replace(col_neso, 'L.pango_Barry', 'grey')
col_neso<-replace(col_neso, 'L.huak', 'grey')
col_neso<-replace(col_neso, 'L.hukare_Barry', 'grey')
#shifting some colours around to make it easier to see
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_neso)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA leio nesi")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#having a look at sams unknowns
col_one<-as.character(ifelse(unique(PCaxis$bee_species_baz) == 'L.unknown','red','grey'))
#col_one
names(col_one)<-as.character(unique(PCaxis$bee_species_baz))
#plot
ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species_baz))+
  theme_bw()+
  scale_fill_manual(values=col_one)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA uknowns")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')+ 
  geom_text_repel(data=subset(PCaxis, bee_species_baz == 'L.unknown'),
            aes(x=Comp1,y=Comp2,label=ID))

#shape extremes in PCA
#bring in links file for plotting
bee_links<-read.table('C:/Users/hrlexd/Dropbox/PlantAndFood (1)/NativeBees/Bee_pictures/Sam_april/Forewing/links_all_landmarks_remove1_sam_hind.txt',sep=',',header=F)
#find mean shape
M<-mshape(beesgpa$coord)
#find extremes of PC1 (max/min)
PC<-PCA$x[,1]
preds<-shape.predictor(beesgpa$coords,x=PC,Intercept=F,pred1=min(PC),pred2=max(PC))
#PC1 min
plotRefToTarget(M,preds$pred1,links=bee_links)
#PC1 max
plotRefToTarget(M,preds$pred2,links=bee_links)
PC<-PCA$x[,2]
preds<-shape.predictor(beesgpa$coords,x=PC,Intercept=F,pred1=min(PC),pred2=max(PC))
#PC2 min
plotRefToTarget(M,preds$pred1,links=bee_links)
#PC2 max
plotRefToTarget(M,preds$pred2,links=bee_links)
```

## Within Leioproctus Leioproctus
Just wanted to have a quick look just within leioproctus leiopructus as I notice on the PCA most of the variaiton is just between Neso species and Leio species


```{r gpaleio}
#just looking with in leioproctus (plus monticola)
#fulvesencs and paahua are quite different
#redo procrustes just with Leioproctus leioproctus species (plus monitola)
Sample_names
bees_leioproctus <- bees[,,-c(8,12:16,28:29,35:39,46:50,55:59,75:79)] 
bees_leioproctus <- bees_leioproctus[-1,,] 
beesgpa_leio<-gpagen(bees_leioproctus)
Sample_names_leio<-labels(bees_leioproctus)[[3]]
Sample_names_leio<-gsub('.*\\\\','',Sample_names_leio)
species_list_leio<-gsub('_.*','',Sample_names_leio)
species_list_leio<-gsub('L.Kanap','L.kanap',species_list_leio)
unique(species_list_leio)
plot(beesgpa_leio)
#plot outliers
plotOutliers(beesgpa_leio$coords)
#run PCA analysis
PCA<-gm.prcomp(beesgpa_leio$coords)
#summary(PCA)
#plot(PCA)
#plot with legend
PCaxis<-as.data.frame(PCA$x[,1:2])
PCaxis<-rownames_to_column(PCaxis,var='ID')
PCaxis$bee_species<-gsub('_[0-9].*','',PCaxis$ID)

PC1_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[1,1]*100,digits=2)
PC2_per<-round(as.data.frame(PCA$sdev^2/sum(PCA$sdev^2))[2,1]*100,digits=2)

ggplot(PCaxis,aes(x=Comp1,y=Comp2,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species))+
  theme_bw()+
  scale_fill_manual(values=col_baz)+
  labs(colour='Species',x=paste0('PC1 ',PC1_per,'%'),y=paste0('PC1 ',PC2_per,'%'),title="PCA")+
  geom_hline(yintercept=0, linetype='dotted', col = 'black')+
  geom_vline(xintercept=0, linetype='dotted', col = 'black')
#
```


## Running some tests
Im not familar with geomorph so this is largely taken from the tutorial online

## Relationship between size and shape

```{R geomorphstats}
#check if there is a relationship between size and shape
#just having a look at raw csize
#metal, bolt kanap seem largest in general 
size<-as.data.frame(beesgpa$Csize)
arrange(size,desc(size$`beesgpa$Csize`)) #%>% print(n=nrow(.))
gdf<-geomorph.data.frame(beesgpa,species=species_list_barry)
fit.size<-procD.lm(coords~log(Csize), data = gdf, print.progress=F)
summary(fit.size)
#is a significant relationship between size and shape
#visiualise
#plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='PredLine',col=as.factor(species_list),pch=19)
#plot with legend etc
p<-plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='PredLine')
p<-cbind(as.data.frame(p$plot_args$x),as.data.frame(p$plot_args$y))
names(p)[names(p) == 'p$plot_args$x'] <- 'PredLine'
names(p)[names(p) == 'p$plot_args$y'] <- 'PC1'
p<-rownames_to_column(p,var='ID')
p$bee_species<-gsub('_[0-9].*','',p$ID)

ggplot(p,aes(x=PredLine,y=PC1,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species))+
  theme_bw()+
  scale_fill_manual(values=col_baz)+
  labs(colour='Species',x='Predictor',y='PC1 for fitted values')

#same thing here:
#plot(fit.size,type='regression',reg.type="PredLine", predictor =log(gdf$Csize))
#first principal component scores fitted values - PredLine

#regression scores-standardised regression shapes 
#plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='RegScore',col=as.factor(species_list),pch=19)
#plot with legend etc
p<-plotAllometry(fit.size,size=gdf$Csize, logsz=TRUE,method='RegScore')
p<-cbind(as.data.frame(p$plot_args$x),as.data.frame(p$plot_args$y))
names(p)[names(p) == 'p$plot_args$x'] <- 'PredLine'
names(p)[names(p) == 'V1'] <- 'RegScore'
p<-rownames_to_column(p,var='ID')
p$bee_species<-gsub('_[0-9].*','',p$ID)

ggplot(p,aes(x=PredLine,y=RegScore,colour = bee_species))+
  geom_point(size=3,colour="grey",shape=21,aes(fill=bee_species))+
  theme_bw()+
  scale_fill_manual(values=col_baz)+
  labs(colour='Species',x='Predictor',y='Regression Score')

#partial least squares for relationship between shape and size
PLS<-two.b.pls(log(gdf$Csize),gdf$coords,print.progress=F)
PLS
#
```

## Relationship between size, shape and species
using Procrustes ANOVA/regression. Not convienced that I have this right and that this is the best way in regards to the posthoc pairwise tests. There is another function which I havent really explored that might be better for this. 

```{R geomorphstats2}
#look at the influence of shape, size and species
fit.common<-procD.lm(coords~log(Csize)+species,data=gdf)
summary(fit.common)
PW<-pairwise(fit.common,groups=species_list_barry)
summary(PW)
#relationship between size and species
fit1<-procD.lm(log(gdf$Csize)~species_list_barry)
summary(fit1)
PW<-pairwise(fit1,groups=species_list_barry)
summary(PW)
#relationship between shape and species
fit2<-procD.lm(gdf$coords~species_list_barry)
summary(fit2)
PW<-pairwise(fit2,groups=species_list_barry)
summary(PW)
#interaction model - shape differ between species, while accounting for shape co varying with size
#not convinced I have this model right, I think its an over simplification
fit3<-procD.lm(gdf$coords ~log(gdf$Csize)*species_list_barry)
summary(fit3)
#posthoc test between groups
PW<-pairwise(fit3,groups=species_list_barry)
summary(PW)
#
```






